Automatic Matching of Bullet Lands

### 2015-12-24 imaging-paper.Rnw
proof-read up to 1111

@@

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats.csv")
names(bstats)[9:10] <- c("#matches", "#non-matches")

@@

\subsection{Algorithm}
<<twolands100, cache=TRUE,  echo=FALSE>>=
images <- file.path(datadir, dir(datadir))

@@

\subsection{Horizontal alignment}
A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.  This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

@@

\section{Evaluation}
\hh{In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
}
\hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land. For this, we assess cross-correlation of signatures at cross sections 25$\mu m$ apart. For a stable region, we require a minimum of 0.9 for the cross correlation. Four lands from different bullets are flagged as problematic in this respect. A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface, also known as `tank rash'.}
%
\begin{figure}
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Br6 Bullet 2-1}
%' <<echo=FALSE>>=
%' b1 <- read.x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
%' image(b1$surface.matrix)
%' @
\includegraphics[width=\textwidth]{images/br6-2-1-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Br9 Bullet 2-4}
\includegraphics[width=\textwidth]{images/br9-2-4-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unkn Bullet B-2}
\includegraphics[width=\textwidth]{images/b-2-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unkn Bullet Q-4}
\includegraphics[width=\textwidth]{images/q-4-grey.png}
\end{subfigure}
\caption{\label{fig:fourflags}\hh{Images of the four lands that got flagged during the quality assessment. All of them show scratch marks across the striation marks from the barrel. They will be excluded from the remainder of the analysis.}}
\end{figure}
%
\hh{We exclude these four lands from the further matching considerations and 
 run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e. we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons. Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM). }
 When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 known lands. Two lands show a high CMS. These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and c. 
 Unfortunately, not all results are as clear cut. 
It might not be reasonable to assume that we can match all lands, but the idea is to try to maximize the number of matches to get an overview of what we can reasonably expect from an automated match.

@@

qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
@
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which.max(cmsdist)]]  
#res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

@@

@
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land with the second top matching CMS.}
<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which(cmsdist==14)]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

@@

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 42 pairs of lands with at least thirteen CMS in common are matches. There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches. On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}). Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}. Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}. Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae. Visually we can easily tell these two lands do not match well.
%
\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{\textwidth}\centering
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[b]{\textwidth}\centering
\caption{Conditional barchart \citep{hummel}: heights show probability of match/non-match given a specific CMS. All land-to-land comparisons with at least 13 CMS are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='.5\\textwidth'>>=
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer(palette="Paired") +
  xlab("maximum CMS")

@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
<<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
x <- seq(0,29, by=0.03)
km <- sm.density(bstats$CMS[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$CMS[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

densities <- data.frame(
  match = rep(c("KM", "KNM"), each=(length(x)+2)),
  x = rep(c(0, x, 29), times = 2),
  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
               data=densities, alpha=0.5) +
xlab("maximum CMS") +
  scale_fill_brewer("", palette="Paired") +
  ylab("Density estimates") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)

# ggplot(data=bstats) + 
#     geom_density(aes(x=cms, fill=match), alpha=0.5) +
#   xlab("maximum CMS") +
#   scale_fill_brewer("", palette="Paired") +
#   theme_bw() + 
#   theme(legend.position=c(1,1), legend.justification=c(1,1),
#         legend.background = element_rect(fill=alpha('white', 0.4)))

@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

\begin{figure}[hbtp]
  \centering
<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.5\\textwidth', warning=FALSE>>=
load("data-25-25/unkn47.RData")
res <- reslist[[106]]
res$bullets$bullet <- scrubPath(res$bullets$bullet)

@@

For smaller numbers of CMS, the percentage of false positives quickly increases. However, if we take other features of the image into account, we can increase the number of correct matches considerably: a high amount of cross-correlation between the two cross sections (see Figure~\ref{fig:ccfs-match}) is indicative for a match; e.g.\  in the Hamby study only known matches have a cross-correlation of 0.75 or higher. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:D}Relationship between maximal CMS and distances between cross sections.}
<<D, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
ggplot(data=bstats) + 
    geom_jitter(aes(x=factor(CMS), y=D, colour=match), alpha=0.5) +
   facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
  ylab("Distance between cross sections") +
  scale_colour_brewer(palette="Paired") +
  theme_bw() + 
  theme(legend.position="none") +
    xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=D, fill=match), alpha=0.5) +
  xlab("Distances between cross sections") +
  scale_fill_brewer("", palette="Paired") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4)))

x <- seq(0,20, by=0.02)
km <- sm.density(bstats$D[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$D[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

densities <- data.frame(
  match = rep(c("KM", "KNM"), each=(length(x)+2)),
  x = rep(c(0, x, 29), times = 2),
  estimate = c(0, km$estimate, 0, 0, knm$estimate,0))

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
               data=densities, alpha=0.5) +
  xlab("Distances between cross sections") +
  scale_fill_brewer("", palette="Paired") +
  ylab("Density estimates") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe)

@
\end{subfigure}
\caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
\end{figure}

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[t]{\textwidth}\centering
\caption{\label{fig:ccfs-match}Relationship between maximal CMS and cross-correlation between cross sections.}
<<ccfs-match, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
ggplot(data=bstats) + theme_bw() + 
  geom_jitter(aes(x=factor(CMS), y=ccf, colour=match), alpha=.5) + 
  facet_wrap(~match, labeller="label_both") + 
    scale_colour_brewer(palette="Paired") +
  xlab("maximum CMS") + ylab("Cross-correlation") +
  theme(legend.position="none")
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
<<ccf-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
ggplot(data=bstats) + 
    geom_density(aes(x=ccf, fill=match), alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired") +
  theme_bw() + 
  theme(legend.position=c(1,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) 
@
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Ratio of probabilities of $P(ccf > x \mid KM) / P(ccf > x \mid KNM)$.}
<<ccf-ratio, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
x <- seq(0,1, by=0.001)
library(sm)
km <- sm.density(bstats$ccf[bstats$match], display="none", eval.points=x)
knm <- sm.density(bstats$ccf[!bstats$match], display="none", eval.points=x)
dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)

densities <- data.frame(
  x = rep(c(0, x, 1), times = 2),
  estimate = c(0, km$estimate, 0, 0, knm$estimate,0),
  match = rep(c("KM", "KNM"), each=1003))

ggplot() +
  geom_polygon(aes(x=x, y=estimate, fill=match), colour="black",
               data=densities, alpha=0.5) +
  xlab("Cross-correlation between cross sections") +
  scale_fill_brewer("", palette="Paired") +
#  scale_colour_brewer("", palette="Paired", labels=c("KNM", "KM")) +
  theme_bw() + 
  ylab("Density estimate") +
  theme(legend.position=c(.975,1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) +
  geom_line(aes(x, pmin(5, kmtail/knmtail)), data=dframe) 
@
\end{subfigure}
\caption{A large cross-correlation between cross sections is indicative of matches between lands.}
\end{figure}

>>>>

\renewcommand{\topfraction}{0.99}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.99}	% max fraction of floats at bottom
% parameters for text pages
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}     % 2 may work better
\setcounter{dbltopnumber}{2}    % for 2-column pages
\renewcommand{\dbltopfraction}{0.9}	% fit big float above 2-col. text
\renewcommand{\textfraction}{0.07}	% allow minimal text w. figs
%   Parameters for FLOAT pages (not text pages):
\renewcommand{\floatpagefraction}{0.7}	% require fuller float pages
% N.B.: floatpagefraction MUST be less than topfraction !!
\renewcommand{\dblfloatpagefraction}{0.7}	% require fuller float pages

@@

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats.csv")

@@

\subsection{Algorithm}\label{sec:algorithm}
<<twolands100, cache=TRUE,  echo=FALSE>>=
images <- file.path(datadir, dir(datadir))

@@

\subsection{Horizontal alignment}
A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.  This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("Matches", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

@@

\section{Evaluation}
\hh{In order to get a better understanding of  how  the matching algorithm works in known matches and non-matches, we investigate its performance in a test scenario based on the James Hamby study.
}
\hh{As a first step, we automatically assess the quality of each of the lands by  checking that we can identify a stable region on each land. For this, we assess cross-correlation of signatures at cross sections 25$\mu m$ apart. For a stable region, we require a minimum of 0.9 for the cross correlation. Four lands from different bullets are flagged as problematic in this respect. A visual inspection (see Figure~\ref{fig:fourflags}) shows that each one of these lands shows scratch marks across the surface, also known as `tank rash'.}
%
\begin{figure}
  \centering
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Br6 Bullet 2-1}
%' <<echo=FALSE>>=
%' b1 <- read.x3p(file.path(datadir, "Br6 Bullet 2-1.x3p"))
%' image(b1$surface.matrix)
%' @
\includegraphics[width=\textwidth]{images/br6-2-1-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Br9 Bullet 2-4}
\includegraphics[width=\textwidth]{images/br9-2-4-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unkn Bullet B-2}
\includegraphics[width=\textwidth]{images/b-2-grey.png}
\end{subfigure}
\begin{subfigure}[t]{.49\textwidth}\centering
\caption{Unkn Bullet Q-4}
\includegraphics[width=\textwidth]{images/q-4-grey.png}
\end{subfigure}
\caption{\label{fig:fourflags}\hh{Images of the four lands that got flagged during the quality assessment. All of them show scratch marks across the striation marks from the barrel. They will be excluded from the remainder of the analysis.}}
\end{figure}
%
\hh{We exclude these four lands from the further matching considerations and 
 run all remaining lands from the unknown bullets against all remaining lands from known bullets for matches, i.e. we are running $15 \times 6 -2 = 90 - 2 = 88$ lands from unkown bullets against $2 \times 10 \times 6 -2 = 120 - 2 = 118$ lands from known bullets, yielding a total of $10,384$ land-to-land comparisons. Out of these comparisons, there are 172 known matches (KM), while the rest are known non-matches (KNM). }
 When things go perfectly, they look like the results in Figure~\ref{fig:hamby-perfect}: Figure~\ref{fig:hamby-perfect}a shows the distribution of the number of maximum consecutive matching striae between land C-3 and all 118 known lands. Two lands show a high CMS. These correspond to the known matches with C-3, shown in Figures~\ref{fig:hamby-perfect}b and c. 
 Unfortunately, not all results are as clear cut. 
It might not be reasonable to assume that we can match all lands, but the idea is to try to maximize the number of matches to get an overview of what we might be able to expect from an automated match.

@@

qplot(cmsdist, geom="bar") + theme_bw() + xlab("Number of CMS")
@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Overlaid signatures of C-3 and the land with the top matching CMS.}
<<top, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which.max(cmsdist)]]  
#res <- reslist[[which.max(cmsdist[-which.max(cmsdist)])]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

@@

@
\end{subfigure}
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Top 2 match with C-3 based on CMS.}
<<top2, echo=FALSE, fig.width=7, fig.height=3.25, out.width='\\textwidth', warning = FALSE>>=
res <- reslist[[which(cmsdist==14)]]  # number 2
res$bullets$bullet <- scrubPath(res$bullets$bullet)

@@

Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 42 pairs of lands with at least thirteen CMS in common are matches. 
%
\begin{figure}[hbtp]
  \centering
%' \begin{subfigure}[b]{\textwidth}\centering
%' \caption{}
\begin{minipage}[t]{.47\textwidth}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\textwidth'>>=
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
%\end{subfigure}
%\begin{subfigure}[b]{\textwidth}\centering
\end{minipage}
\begin{minipage}[t]{.52\textwidth}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=4, out.width='\\textwidth'>>=
bstats$km <- c( "KNM", "KM")[as.numeric(bstats$match)+1]
ggplot(data=subset(bstats, !flagged)) + 
  geom_bar(aes(x=factor(CMS), fill=km), position="fill") +
  theme(legend.position="bottom") + 
  scale_fill_brewer("", palette="Paired") +
  xlab("maximum CMS")

@
\end{minipage}
%' \end{subfigure}
%' \begin{subfigure}[t]{\textwidth}\centering
%' \caption{\label{fig:cms-dens}Densities of maximum CMS for Known Matches (KM) and Known Non Matches (KNM).}
%' <<cms-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.6\\textwidth'>>=
%' x <- seq(0,29, by=0.03)
%' km <- sm.density(bstats$CMS[bstats$match],h=2, model="Normal", display="none", eval.points=x)
%' knm <- sm.density(bstats$CMS[!bstats$match],h=1, display="none", eval.points=x)
%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
%' 
%' densities <- data.frame(
%'   match = rep(c("KM", "KNM"), each=(length(x)+2)),
%'   x = rep(c(0, x, 29), times = 2),
%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
%' 
%' ggplot() +
%'   geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
%'                data=densities, alpha=0.5) +
%' xlab("maximum CMS") +
%'   scale_fill_brewer("", palette="Paired") +
%'   ylab("Density estimates") +
%'   theme_bw() + 
%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' #  geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe) 
%' 
%' # ggplot(data=bstats) + 
%' #     geom_density(aes(x=cms, fill=match), alpha=0.5) +
%' #   xlab("maximum CMS") +
%' #   scale_fill_brewer("", palette="Paired") +
%' #   theme_bw() + 
%' #   theme(legend.position=c(1,1), legend.justification=c(1,1),
%' #         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' 
%' @
%' \end{subfigure}
\caption{\label{fig:cms}Distribution of maximal CMS (left). Conditional barchart \citep{hummel} on the right: heights show probability of match/non-match given a specific CMS. All land-to-land comparisons with at least 13 CMS are matches.}
\end{figure}
%
There are two things that should be noted at this point: the automated algorithm finds a relatively high number of CMS even for non-matches. On average, there are \Sexpr{mean(bstats$cms[!bstats$match])} maximal CMS between known non-matches (with a standard deviation of \Sexpr{sd(bstats$cms[!bstats$match])}). Known matches share on average \Sexpr{mean(bstats$cms[bstats$match])} maximal CMS, with a standard deviation of \Sexpr{sd(bstats$cms[bstats$match])}. Secondly, a large number of maximal CMS by itself is not indicative of a match, as was previously pointed out by \citet{miller:1998}. Figure~\ref{fig:mismatch} shows a known mismatch, between two lands that share twelve consecutively matched striae. Visually we can easily tell that these two lands do not match well.

\begin{figure}[hbtp]
  \centering
<<strange-res, echo=FALSE, fig.width=7, fig.height=3.25, out.width='.65\\textwidth', warning=FALSE>>=
load("data-25-25/unkn47.RData")
res <- reslist[[106]]
res$bullets$bullet <- scrubPath(res$bullets$bullet)

@@

For smaller numbers of CMS, the percentage of false positives quickly increases. However, if we take other features of the image into account, we can increase the number of correct matches considerably: Figure~\ref{fig:densities} gives an overview of the densities of all of the features derived earlier, distinguishing between known matches (KM) and known non-matches (KNM). The densities of almost all of the features show strong differences between matches and non matches, e.g\ a high amount of cross-correlation between two signatures is indicative for a match --  in the Hamby study only known matches have a cross-correlation of 0.75 or higher. There are 97 land-to-land comparisons with a cross-correlation of above 0.75.

\begin{figure}[hbtp]
  \centering
<<density-overview,echo=FALSE, warning=FALSE, fig.width=12, fig.height=7, out.width='\\textwidth'>>=
features <- c("CMS", "CNMS",  "num.matches", "num.nonmatches", "D", "S", "ccf")
densities <- plyr::ldply(features, function(x) {
  xr <- range(bstats[,x])
  xx <- seq(xr[1], xr[2], length.out=500)

  densKM <- sm.density(bstats[,x][bstats$match], display="none", 
                       eval.points=xx, weights=NA, method="normal")
  densKNM <- sm.density(bstats[,x][!bstats$match], display="none", 
                       eval.points=xx, weights=NA, method="normal")
  dframe <- data.frame(var=x, x = xx, KM=densKM$estimate, KNM = densKNM$estimate)
  dframe <- rbind(data.frame(var=x, x = xr[1], KM=0, KNM=0),
        dframe, data.frame(var=x, x = xr[2], KM=0, KNM=0))
  dframe$order <- 1:nrow(dframe)
  dframe
}) 
library(reshape2)
dm <- melt(densities, measure.var=c("KM", "KNM"))
dm$var <- factor(dm$var, levels=features)
levels(dm$var)[3:4] <- c("#matches", "#non-matches")

qplot(x, value, group=variable, geom="polygon", data=dm, fill=variable, 
      alpha=I(0.6), colour=I("grey20")) + 
  facet_wrap(~var, ncol=4, scales="free") +
  theme_bw() + ylab("density") + xlab("") +
  scale_fill_brewer("Matches", palette="Paired") +
  theme(legend.position=c(1,0), legend.justification = c("right", "bottom"),
        legend.background = element_rect(colour="grey60"),
        plot.margin = unit(c(0,0,0,0), unit="cm"))
@
\caption{\label{fig:densities}Overview of all the marginal densities for features extracted in section~\ref{sec:algorithm}.}
\end{figure}

%' \begin{figure}[hbtp]
%'   \centering
%' \begin{subfigure}[t]{\textwidth}\centering
%' \caption{\label{fig:D}Relationship between maximal CMS and distances between cross sections.}
%' <<D, echo=FALSE, fig.width=8, fig.height=4, out.width='.5\\textwidth'>>=
%' ggplot(data=bstats) + 
%'     geom_jitter(aes(x=factor(CMS), y=D, colour=match), alpha=0.5) +
%'    facet_wrap(~match, labeller="label_both") + xlab("maximum CMS") + 
%'   ylab("Distance between cross sections") +
%'   scale_colour_brewer(palette="Paired") +
%'   theme_bw() + 
%'   theme(legend.position="none") +
%'     xlab("maximum CMS")
%' @
%' \end{subfigure}
%' \begin{subfigure}[t]{\textwidth}\centering
%' \caption{Densities of distances between cross sections for Known Matches (KM) and Known Non Matches (KNM).}
%' <<crosscut-densities, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='.5\\textwidth'>>=
%' ggplot(data=bstats) + 
%'     geom_density(aes(x=D, fill=match), alpha=0.5) +
%'   xlab("Distances between cross sections") +
%'   scale_fill_brewer("", palette="Paired") +
%'   theme_bw() + 
%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
%'         legend.background = element_rect(fill=alpha('white', 0.4))) 
%' 
%' 
%' 
%' x <- seq(0,20, by=0.02)
%' km <- sm.density(bstats$D[bstats$match], display="none", eval.points=x)
%' knm <- sm.density(bstats$D[!bstats$match], display="none", eval.points=x)
%' dframe <- data.frame(x = x, km=km$estimate, knm=knm$estimate)
%' dframe$kmtail <- rev(cumsum(rev(dframe$km)))/sum(dframe$km)
%' dframe$knmtail <- rev(cumsum(rev(dframe$knm)))/sum(dframe$knm)
%' 
%' densities <- data.frame(
%'   match = rep(c("KM", "KNM"), each=(length(x)+2)),
%'   x = rep(c(0, x, 29), times = 2),
%'   estimate = c(0, km$estimate, 0, 0, knm$estimate,0))
%' 
%' ggplot() +
%'   geom_polygon(aes(x=x, y=estimate, fill=match, group=match), colour="black",
%'                data=densities, alpha=0.5) +
%'   xlab("Distances between cross sections") +
%'   scale_fill_brewer("", palette="Paired") +
%'   ylab("Density estimates") +
%'   theme_bw() + 
%'   theme(legend.position=c(1,1), legend.justification=c(1,1),
%'         legend.background = element_rect(fill=alpha('white', 0.4))) +
%'   geom_line(aes(x=x, y=pmin(0.2, kmtail/knmtail/10)), data=dframe) 
%' 
%' @
%' \end{subfigure}
%' \caption{Relationship between distances between cross sections. A small distance between cross sections is indicative of matches between lands.}
%' \end{figure}
